---
title: 'Homework #2'
author: "Mallory Lai"
date: "October 30, 2016"
output: pdf_document
---
```{r, message = F, warning=F}
library(markovchain)
library(igraph)
library(knitr)
```


1.
p~01~ = 1; 100% probability of going from 0 to 1. Therefore, p~00~ = 0 for all n.  
p~1,i+1~ = 1; 100% probability of going to either i+1 or i-1 from i. No other possibilities!  
$\frac{p~i,i+1~}{p~i,i-1}= (\frac{i+1}{i})^2$ Since the numerator is greater 
than the denominator, the chain has a tendency to move to the right. 
```{r}
# Create graph using igraph.
# Set seed to ensure graph doesn't change. 
set.seed(3)
g1 <- graph(c('i','i+1', 'i','i-1')) %>% 
  set_edge_attr("label", value = c('numerator', 'denominator'))
plot(g1, vertex.size = 50, vertex.color = "green", edge.arrow.size = .5)

```

2. Setting up the problem:

```{r}
# Create transition matrix
m <- matrix(c(0, .5, .5, 0, 0,
              0, .25, .25, .5, 0,
              0, .5, .5, 0, 0, 
              0, 0, 0, .25, .75,
              0, 0, 0, .5, .5), 
            ncol = 5, byrow = T)

# Create a discrete time Markov Chain using the package "markovchain".
MC1 <- as(m, "markovchain")

# Set seed to prevent graph from changing. 
set.seed(1)
# Graph MC1. 
plot(MC1, vertex.size = 20)
```

a) Classify the states. 
```{r}
# Examine summary of MC1
summary(MC1)
```
4 and 5 are both a closed and recurrent class. 1 is a transient class that can never be returned to once left. 2 and 3 are transient as well. 

b) Determine all the invariant probabilities. 
```{r}
steadyStates(MC1)
```
The invariant probabilities are $\pi$ = {0, 0, 0, 0.4, 0.6}.

c) What is the probability to be in {4,5} starting from 1. 
```{r}
MC1[1, c(4,5)]
```

There is a 0% probability of moving to {4,5} from 1. The chain can only move to
2 or 3 from 1. In terms of a steady state solution, however, there is a 100% of
being in {4,5} since all other states are transient and {4,5} is closed. 

3.

a.

```{r}
# Create graph using igraph.
# Set seed to ensure graph doesn't change. 
set.seed(2)
g4 <- graph(c(1,2, 2,3, 3,1, 3,2, 2,1, 1,3)) %>%
  set_edge_attr("label", value = c('1-p', '1-p', '1-p', 'p', 'p', 'p'))
plot(g4, vertex.size = 50, vertex.color = "green", 
edge.arrow.size = .5)

```
  
b.
$$P =
 \begin{pmatrix}
  0 & 1-p & p \\
  p & 0 & 1-p \\
  1-p & p & 0
 \end{pmatrix}$$
 
c. Since the R package cannot handle variables, only numerical values, I will
use 10 random values between 1 and 10 to show that the invariant probability 
does not depend on the value of p. 
```{r}
# Create a vector of probabilities for p to take on. 
p <- runif(10)

# Initialize an empty matrix to hold invariant probabilities for each p. 
invariantProbabilities <- matrix(NA, nrow = length(p), ncol = 3)

# Create a Markov Chain and calculate invariant probabilities for each p.
for(i in 1:length(p))
{
  
# Create transition matrix.  
  m <- matrix(c(0, 1-p[i], p[i],
              p[i], 0, 1-p[i],
              1-p[i], p[i], 0), 
            ncol = 3, byrow = T)

# Create a discrete time Markov Chain using the package "markovchain".
MC <- as(m, "markovchain")

# Determine invariant probabilities.
invariantProbabilities[i, ] <- steadyStates(MC)

}

# Assign rownames that match with p value. 
rownames(invariantProbabilities) <- round(p, digits = 2)
# Assign column names that match with vertices. 
colnames(invariantProbabilities) <- c(1, 2, 3)

# Show invariant probabilities for all values of p. 
kable(invariantProbabilities)
```

The invariant probabilities are $\pi$ = {1/3, 1/3, 1/3}, regardless of the 
starting value of p. 
  
d. $$P(X_{n} = 1, X_{n + 1} = 2) = \frac{1-p}{3}$$  
  $$P(X_{n} = 2, X_{n + 1} = 1) = \frac{p}{3}$$
