---
title: 'Homework #2'
author: "Mallory Lai"
date: "October 30, 2016"
output: html_document
---


1.


2. Setting up the problem:
```{r, message = F, warning=F}
library(markovchain)
library(igraph)
```

```{r}
# Create transition matrix
m <- matrix(c(0, .5, .5, 0, 0,
              0, .25, .25, .5, 0,
              0, .5, .5, 0, 0, 
              0, 0, 0, .25, .75,
              0, 0, 0, .5, .5), 
            ncol = 5, byrow = T)

# Create a discrete time Markov Chain using the package "markovchain".
MC1 <- as(m, "markovchain")

# Graph MC1. 
plot(MC1)
```

a) Classify the states. 
```{r}
# Examine summary of MC1
summary(MC1)
```

b) Determine all the invariant probabilities. 
```{r}
steadyStates(MC1)
```


c) What is the probability to be in {4,5} starting from 1. 
```{r}
MC1[1, c(4,5)]
```

There is a 0% probability of moving to {4,5} from 1. The chain can only move to
2 or 3 from 1. In terms of a steady state solution, however, there is a 100% of
being in {4,5} since all other states are transient and {4,5} is closed. 

3.

a.

```{r}
# Create vertices of triangle.
nodes <- c(1, 2, 3)

# State probabilities between vertices. 
links <- c("1-p", "p")



```

b.

```{r}
# Create transition matrix.  
  (m <- matrix(c(0, "1-p", "p",
              "p", 0, "1-p",
              "1-p", "p", 0), 
            ncol = 3, byrow = T))
```

c. Since the R package cannot handle variables, only numerical values, I will
use 10 random values between 1 and 10 to show that the invariant probability 
does not depend on the value of p. 
```{r}
# Create a vector of probabilities for p to take on. 
p <- runif(10)

# Initialize an empty matrix to hold invariant probabilities for each p. 
invariantProbabilities <- matrix(NA, nrow = length(p), ncol = 3)

# Create a Markov Chain and calculate invariant probabilities for each p.
for(i in 1:length(p))
{
  
# Create transition matrix.  
  m <- matrix(c(0, 1-p[i], p[i],
              p[i], 0, 1-p[i],
              1-p[i], p[i], 0), 
            ncol = 3, byrow = T)

# Create a discrete time Markov Chain using the package "markovchain".
MC <- as(m, "markovchain")

# Determine invariant probabilities.
invariantProbabilities[i, ] <- steadyStates(MC)

}

invariantProbabilities
```

The invariant probabilities are \Pi = {1/3, 1/3, 1/3}.

```{r}

```




